---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ghinspectoR

<!-- badges: start -->
<!-- badges: end -->

This package originated from work done for a GitHub dashboard which was set up 
to check over multiple repositories within an organisation. 
It started off with a focus on code owners, stale branches and techdebt issues, 
and with the help of Copilot, has been extended to a series of functions that
`get_` list data from the GitHub API and then `tidy_` it for R user into a 
more familiar data frame (tibble) view.

The package uses authentication with GitHub which relies upon the setting up of
an authentication app across the organisation. This has been detailed in a 
[blog](https://the-strategy-unit.github.io/data_science/blogs/posts/2025-04-30_using-github-apps-for-authentication/)
on The Strategy Unit Data Science website.

## Installation

You can install the development version of ghinspectoR:

``` r
# install.packages("pak")
pak::pak("The-Strategy-Unit/ghinspectoR")
```
# Github API set up

These functions require a token to be set up on GitHub to connect a GitHub App 
which will be located in the organisation's page then Settings > GitHub Apps.

The link will look something similar to this with `{organisation-name}` replaced
with the actual name.

`https://github.com/organizations/{organisation-name}/settings/installations`

> [!IMPORTANT]  
> GitHub Apps can only be set up and maintained by a GitHub Organization Owner. 
> This includes generating of private keys which has details below. 
> Two factor authentication will also be required to access this area.

Adding an account that is not an Owner can be done through Settings 
(at the Organisational level) > GitHub Apps > Configure (the specific App) > App Settings > App managers

## Local set up of .pem file

Once a GitHub App is set up a connection between that and the local version 
of the repository on a computer needs to be set up.
This is done using a `.pem` file which is downloaded from GitHub located in the
[Developer Settings](https://docs.github.com/en/apps/creating-github-apps/authenticating-with-a-github-app/managing-private-keys-for-github-apps)

If multiple people have access each will have their own `.pem` file.

The private key file will be downloaded with this format of name:
`<name of app>.<date>.private-key.pem` and needs to be moved from the 
Downloads folder to somewhere that will be easily accessible to the Project.
This is preferably outside the Project itself but `*.pem` can also be 
added to a `.gitignore` to avoid accidental committing.

## Setting up local credentials using [{keyring}](https://keyring.r-lib.org/) package

The {keyring} package is a secure way to give credential access across multiple 
projects. 
Set up with the following:

```{r}
keyring::key_set("GITHUB_APP_ID")
```

Add the App ID found from the GitHub App details. 

Then run:

```{r}
keyring::key_set("GITHUB_APP_PRIVATE_KEY")
```

Putting in the password box the pathway to the `.pem` file but importantly, with
no quotations so like `C:\Users\name\Documents\SECRETS\GITHUB_APP_PRIVATE_KEY.pem`

To view the keyring:

```{r}
keyring::key_get("GITHUB_APP_PRIVATE_KEY")
keyring::key_get("GITHUB_APP_ID")
```

Note that for Windows users the `\` in the file pathways get translated by the 
{keyring} package with `\\` automatically.

## Setting up local credentials using `.Renviron`

In some circumstances users may use a `.Renviron` file to save credentials and 
in these circumstances ensure that the file is in the `.gitignore` to prevent
accidental committing to GitHub.

Opening the `.Renviron` using the {usethis} package:

```
usethis::edit_r_environ(scope = "project")

```

Note this is opens the project `.Renviron` and so will need to be set up for 
any project that requires this connection.
Also note that the {usethis} function gives the link to the file and reminds
you to restart R after any changes in the Console.

In the file add the text:

```
GITHUB_APP_ID=1111111111111111
GITHUB_APP_PRIVATE_KEY="C:\Users\name.lastName\Downloads\GITHUB_APP_PRIVATE_KEY.pem"

```

These need to be updated with real information for the ID and key pathway.
Unlike in the keyring set up, the file pathway in the `.Renviron` requires
quotes around the text.

## Posit Connect set up

On the Posit Connect published page, in the Settings > Vars the variables are
also saved as GITHUB_APP_ID and GITHUB_APP_PRIVATE_KEY.

> [!WARNING]
> Details of variables saved on Posit Connect cannot be viewed, only replaced.

## Further reading

[GitHub Apps for Authentication blog](https://the-strategy-unit.github.io/data_science/blogs/posts/2025-04-30_using-github-apps-for-authentication/)
